Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-16 csf-2017/paper.tex
Fix citations to use IEEE style and to use natbib properly everywhere.

\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{natbib}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

@@

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control-flow
correctness, local-state encapsulation, etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control-flow correctness
and local-state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) and/or rely on techniques like
machine code rewriting
(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a
native stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
randomization~\citep{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

\emph{Capability machines} are a type of processors that
remediate these limitations with a better security model at
the hardware level. They are based on old ideas
(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
but have recently received renewed interest; in particular, the CHERI
project has proposed new ideas and ways of tackling practical
challenges like backwards compatibility and realistic OS
support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities (a kind
of pointers that carry authority). Memory capabilities carry
the authority to read and/or write to a range of memory
locations. There is also a form of \emph{object capabilities}, which represent the
authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g., the M-Machine's enter capabilities or
CHERI's sealed code/data pairs).

@@

In this paper, we make progress towards the goal of efficient and provably
watertight enforcement of encapsulation and control-flow correctness on
capability machines. Specifically, we make the following contributions:
\begin{itemize}
\item Formalize a simple but representative capability machine featuring local
  capabilities, and its operational semantics
  (Section~\ref{sec:capab-mach-with})
\item Define a novel calling convention enforcing control-flow correctness and
  encapsulation of stack frames (Section~\ref{sec:stack-and-return-pointer}). It
  relies solely on local capabilities and does not require OS support (like a
  trusted stack or call/return instructions). It also does not use separate
  per-component stacks but a single, fairly standard, contiguous stack. It
  supports higher-order cross-component calls (e.g., cross-component function
  pointers) and can be efficient assuming only one additional piece of processor
  support: an efficient instruction for clearing a range of memory.
\item Prove a theorem that is the most general and powerful formulation of the
  guarantees offered by a capability machine (a form of capability
  safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
  guarantees offered for local capabilities. It is very general and not tied to
  our calling convention or a specific way of using the system's capabilities.
\item Introduce two novel technical ideas in the unary, step-indexed Kripke
  logical relation used to formulate the above theorem. First, we model the
  naturally continuation-passing-style of assembly code (with return pointers as
  continuations) using a \emph{single} orthogonal closure, rather than the
  existing biorthogonal closure. Secondly, we express the special nature of
  local capabilities using a variant of Dreyer et. al.'s public and private
  future worlds~\citep{Dreyer:jfp12}, used to express well-bracketedness
  guarantees in a lambda calculus. The logical relation and capability safety
  theorem are presented in Section~\ref{sec:logical-relation}.
\item Demonstrate our results by applying them to challenging examples,
  specifically constructed to demonstrate local-state encapsulation and
  control-flow correctness guarantees in the presence of cross-component
  function pointers (Section~\ref{sec:examples}). The examples demonstrate both
  the power of our formulation of capability safety and our calling convention.
\end{itemize}
%\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution. Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}

For reasons of space, some details and all proofs have been omitted;
refer to the technical appendix~\cite{technical_appendix} for those.

@@

\section{A capability machine with local capabilities}
\label{sec:capab-mach-with}
In this paper, we work with a formal capability machine with all the
characteristics of real capability machines, as well as local capabilities much
like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
both the M-Machine~\cite{Carter:1994:HSF:195473.195579} and
CHERI~\cite{Watson2015Cheri}. To avoid uninteresting details, we assume an
infinite address space and unbounded integers.

@@

The $\jmp{}$ instruction updates the program counter to a requested location,
but it is complicated by the presence of \emph{enter capabilities}, modeled
after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
cannot be used to read, write or execute and their address and range cannot be
modified. They can only be used to jump to, but when that happens, their
permission changes to $\exec$. They can be used to represent a kind of closures:
an opaque package containing a piece of code together with local encapsulated
state. Such a package can be built as an enter capability $c =
((\entry,\gl),\start,\addrend,\addr)$ where the range $[\start,\addr-1]$
contains local state (data or capabilities) and $[\addr,\addrend]$ contains
instructions. The package is opaque to an adversary holding $c$ but when $c$ is
jumped to, the instructions can start executing and have access to the local
data through the updated version of $c$ that is then in $\pcreg$\footnote{Note
  that the capability in the $\pcreg$-register can be used to load from.}.
% \lau{As we
%   talked about what is described here, is not how we make closures in the TR, so
%   we should consider whether we should refrain from using that word here. }

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\cite{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we take care to clear the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as all
capabilities with write-local authority are themselves local and the stack is
cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

@@

In order to have a reasonably realistic system, we use a simple model of
linking, so programs can have access to other programs through a linking table.
We also assume malloc to be part of the trusted computing base satisfying a
certain specification. Malloc and linking tables are described further in the
next section, but we refer to the technical appendix~\cite{technical_appendix}
for full details.

@@

The technical appendix~\cite{technical_appendix} contains detailed descriptions
of all the macros.

@@

\subsection{Worlds}
\begin{figure}
  \centering
  \includegraphics{w11}
  \caption{An illustration of local/global capabilities and temporary/permanent
    regions. The table represents a memory and the colored fields are regions
    governing a memory range. Global capabilities cannot depend on temporary
    regions.}
  \label{fig:cap-world}
\end{figure}
% \dominique{Perhaps we could first try to convey the intuition of a permanent
%   world with a temporary overlay, with global capabilities ignoring the overlay.
%   Perhaps we could add a depiction of the world to Figure 5?}
%\todo[inline]{Find a good place to describe monotonicity of
%interpretation function.}
% \lau{this may not be an important enough point to include it.}
A world is a finite map from region names, modeled as natural numbers, to regions.
We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
%$\perma$, $\temp$, and $\revoked$. The
%$\perma$ and $\temp$ 
Each permanent and temporary region contains a state transition system, with
public and private transitions, to describe a protocol for the region's memory,
similarly to what has been used in logical relations for high-level
languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
imposed by permanent regions stay in place indefinitely. Any capability, local
or global, can depend on these protocols. Protocols imposed by temporary regions
can be revoked at certain points during execution, on the condition that we
clear local capabilities when this happens. This means that local capabilities
can safely depend on the protocols imposed by temporary regions, but global
capabilities cannot, since a global capability may outlive a temporary region
that is revoked. This is illustrated in \figurename~\ref{fig:cap-world}.

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{
    \begin{multlined}
      \npairP{(\var{perms}, \\
        \start,\addrend)}
    \end{multlined}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
}
%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation now defines semantically when values, terms, and
configurations can safely be used or executed by the adversary, i.e. without the
risk that they will be able to use it to break protocols in the world. The
definition of the logical relation is found in \figurename{}s~\ref{fig:logrel}
and~\ref{fig:perm-conds} and we provide some explanations in the following
paragraphs. We omit some definitions and explain them only verbally, but precise
definitions can be found in the technical appendix~\cite{technical_appendix}.

The \emph{observation relation} $\observations$ defines configurations that we
consider safe. Roughly speaking, a configuration is safe if whenever it halts,
there is a private future world that the resulting memory satisfies. Notice that
failing is considered safe behavior. In fact, the machine often resorts to
failing when an unauthorized access is attempted, like loading from a capability
without read permission. This is similar to \cite{Devriese:2016ObjCap}'s logical
relation for an untyped language, but unlike typical logical relations
for typed languages, which require that programs do not fail.

@@

\section{Examples}
\label{sec:examples}
In this section, we demonstrate how our formalization of capability safety
allows us to prove local-state encapsulation and control-flow correctness
properties for challenging program examples. The security measures of
Section~\ref{sec:stack-and-return-pointer} are deployed to ensure these
properties. Since we are dealing with assembly language, there are many details
to the formal treatment, and therefore we necessarily have to omit some details
in the lemmas we state here. The interested reader can find all the technical
details in the technical appendix~\cite{technical_appendix}.

@@

\section{Discussion}
\label{sec:discussion}
\noindent\textbf{Calling convention}\\
\emph{Formulating control flow correctness} While we claim that our calling
convention enforces control-flow correctness, we do not prove a general theorem
that shows this, because it is not clear what such a theorem should look like.
Formulations in terms of a control-flow graph, like the one by
\cite{abadi_control-flow_2005}, do not take into account temporal properties,
like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
In fact, our examples show that our logical relation imply a stronger form of
control-flow correctness than such formulations, although this is not made very
explicit. As future work, we consider looking at a more explicit and useful way
to formalize control-flow correctness. The idea would be to define a variant
of our capability machine with call and return instructions and
well-bracketed control flow built-in to the operational semantics, and then
prove that compiling such code to our machine using our calling convention is
fully abstract~\citep{abadi_protection_1998}.

@@

Our logical relation is a unary, step-indexed Kripke logical relation
with recursive
worlds~\cite{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
Fundamental Theorem is similar to theirs and expresses
capability safety of the capability machine. Because we are not interested in
externally observable side-effects (like console output or memory access
traces), we do not require their notion of effect parametricity. Our logical
relation uses several ideas from previous work, like Kripke worlds with regions
containing state transition systems~\citep{Ahmed:popl09}, public/private future
worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

@@

\bibliographystyle{plainnat}
%\bibliographystyle{plain}
\bibliography{IEEEabrv,references}

>>>>

\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage[numbers]{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{multicol}
\lstset{
  basicstyle=\footnotesize\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

@@

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control-flow
correctness, local-state encapsulation, etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control-flow correctness
and local-state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) and/or rely on techniques like
machine code rewriting
(e.g.,~\citet{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
verification (e.g.,~\citet{morrisett_system_1999}), virtual machines with a
native stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
randomization~\citep{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

\emph{Capability machines} are a type of processors that
remediate these limitations with a better security model at
the hardware level. They are based on old ideas
(e.g.~\citet{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
but have recently received renewed interest; in particular, the CHERI
project has proposed new ideas and ways of tackling practical
challenges like backwards compatibility and realistic OS
support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities (a kind
of pointers that carry authority). Memory capabilities carry
the authority to read and/or write to a range of memory
locations. There is also a form of \emph{object capabilities}, which represent the
authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g., the M-Machine's enter capabilities or
CHERI's sealed code/data pairs).

@@

In this paper, we make progress towards the goal of efficient and provably
watertight enforcement of encapsulation and control-flow correctness on
capability machines. Specifically, we make the following contributions:
\begin{itemize}
\item Formalize a simple but representative capability machine featuring local
  capabilities, and its operational semantics
  (Section~\ref{sec:capab-mach-with})
\item Define a novel calling convention enforcing control-flow correctness and
  encapsulation of stack frames (Section~\ref{sec:stack-and-return-pointer}). It
  relies solely on local capabilities and does not require OS support (like a
  trusted stack or call/return instructions). It also does not use separate
  per-component stacks but a single, fairly standard, contiguous stack. It
  supports higher-order cross-component calls (e.g., cross-component function
  pointers) and can be efficient assuming only one additional piece of processor
  support: an efficient instruction for clearing a range of memory.
\item Prove a theorem that is the most general and powerful formulation of the
  guarantees offered by a capability machine (a form of capability
  safety~\citet{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
  guarantees offered for local capabilities. It is very general and not tied to
  our calling convention or a specific way of using the system's capabilities.
\item Introduce two novel technical ideas in the unary, step-indexed Kripke
  logical relation used to formulate the above theorem. First, we model the
  naturally continuation-passing-style of assembly code (with return pointers as
  continuations) using a \emph{single} orthogonal closure, rather than the
  existing biorthogonal closure. Secondly, we express the special nature of
  local capabilities using a variant of Dreyer et. al.'s public and private
  future worlds~\citep{Dreyer:jfp12}, used to express well-bracketedness
  guarantees in a lambda calculus. The logical relation and capability safety
  theorem are presented in Section~\ref{sec:logical-relation}.
\item Demonstrate our results by applying them to challenging examples,
  specifically constructed to demonstrate local-state encapsulation and
  control-flow correctness guarantees in the presence of cross-component
  function pointers (Section~\ref{sec:examples}). The examples demonstrate both
  the power of our formulation of capability safety and our calling convention.
\end{itemize}
%\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution. Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}

For reasons of space, some details and all proofs have been omitted;
refer to the technical appendix~\citep{technical_appendix} for those.

@@

\section{A capability machine with local capabilities}
\label{sec:capab-mach-with}
In this paper, we work with a formal capability machine with all the
characteristics of real capability machines, as well as local capabilities much
like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and
CHERI~\citep{Watson2015Cheri}. To avoid uninteresting details, we assume an
infinite address space and unbounded integers.

@@

The $\jmp{}$ instruction updates the program counter to a requested location,
but it is complicated by the presence of \emph{enter capabilities}, modeled
after the M-Machine's~\citep{Carter:1994:HSF:195473.195579}. Enter capabilities
cannot be used to read, write or execute and their address and range cannot be
modified. They can only be used to jump to, but when that happens, their
permission changes to $\exec$. They can be used to represent a kind of closures:
an opaque package containing a piece of code together with local encapsulated
state. Such a package can be built as an enter capability $c =
((\entry,\gl),\start,\addrend,\addr)$ where the range $[\start,\addr-1]$
contains local state (data or capabilities) and $[\addr,\addrend]$ contains
instructions. The package is opaque to an adversary holding $c$ but when $c$ is
jumped to, the instructions can start executing and have access to the local
data through the updated version of $c$ that is then in $\pcreg$\footnote{Note
  that the capability in the $\pcreg$-register can be used to load from.}.
% \lau{As we
%   talked about what is described here, is not how we make closures in the TR, so
%   we should consider whether we should refrain from using that word here. }

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we take care to clear the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as all
capabilities with write-local authority are themselves local and the stack is
cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

@@

In order to have a reasonably realistic system, we use a simple model of
linking, so programs can have access to other programs through a linking table.
We also assume malloc to be part of the trusted computing base satisfying a
certain specification. Malloc and linking tables are described further in the
next section, but we refer to the technical appendix~\citep{technical_appendix}
for full details.

@@

The technical appendix~\citep{technical_appendix} contains detailed descriptions
of all the macros.

@@

\subsection{Worlds}
\begin{figure}
  \centering
  \includegraphics{w11}
  \caption{An illustration of local/global capabilities and temporary/permanent
    regions. The table represents a memory and the colored fields are regions
    governing a memory range. Global capabilities cannot depend on temporary
    regions.}
  \label{fig:cap-world}
\end{figure}
% \dominique{Perhaps we could first try to convey the intuition of a permanent
%   world with a temporary overlay, with global capabilities ignoring the overlay.
%   Perhaps we could add a depiction of the world to Figure 5?}
%\todo[inline]{Find a good place to describe monotonicity of
%interpretation function.}
% \lau{this may not be an important enough point to include it.}
A world is a finite map from region names, modeled as natural numbers, to regions.
We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
%$\perma$, $\temp$, and $\revoked$. The
%$\perma$ and $\temp$ 
Each permanent and temporary region contains a state transition system, with
public and private transitions, to describe a protocol for the region's memory,
similarly to what has been used in logical relations for high-level
languages~\citep{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
imposed by permanent regions stay in place indefinitely. Any capability, local
or global, can depend on these protocols. Protocols imposed by temporary regions
can be revoked at certain points during execution, on the condition that we
clear local capabilities when this happens. This means that local capabilities
can safely depend on the protocols imposed by temporary regions, but global
capabilities cannot, since a global capability may outlive a temporary region
that is revoked. This is illustrated in \figurename~\ref{fig:cap-world}.

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{
    \begin{multlined}
      \npairP{(\var{perms}, \\
        \start,\addrend)}
    \end{multlined}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
}
%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation now defines semantically when values, terms, and
configurations can safely be used or executed by the adversary, i.e. without the
risk that they will be able to use it to break protocols in the world. The
definition of the logical relation is found in \figurename{}s~\ref{fig:logrel}
and~\ref{fig:perm-conds} and we provide some explanations in the following
paragraphs. We omit some definitions and explain them only verbally, but precise
definitions can be found in the technical appendix~\citep{technical_appendix}.

The \emph{observation relation} $\observations$ defines configurations that we
consider safe. Roughly speaking, a configuration is safe if whenever it halts,
there is a private future world that the resulting memory satisfies. Notice that
failing is considered safe behavior. In fact, the machine often resorts to
failing when an unauthorized access is attempted, like loading from a capability
without read permission. This is similar to \citet{Devriese:2016ObjCap}'s logical
relation for an untyped language, but unlike typical logical relations
for typed languages, which require that programs do not fail.

@@

\section{Examples}
\label{sec:examples}
In this section, we demonstrate how our formalization of capability safety
allows us to prove local-state encapsulation and control-flow correctness
properties for challenging program examples. The security measures of
Section~\ref{sec:stack-and-return-pointer} are deployed to ensure these
properties. Since we are dealing with assembly language, there are many details
to the formal treatment, and therefore we necessarily have to omit some details
in the lemmas we state here. The interested reader can find all the technical
details in the technical appendix~\citep{technical_appendix}.

@@

\section{Discussion}
\label{sec:discussion}
\noindent\textbf{Calling convention}\\
\emph{Formulating control flow correctness} While we claim that our calling
convention enforces control-flow correctness, we do not prove a general theorem
that shows this, because it is not clear what such a theorem should look like.
Formulations in terms of a control-flow graph, like the one by
\citet{abadi_control-flow_2005}, do not take into account temporal properties,
like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
In fact, our examples show that our logical relation imply a stronger form of
control-flow correctness than such formulations, although this is not made very
explicit. As future work, we consider looking at a more explicit and useful way
to formalize control-flow correctness. The idea would be to define a variant
of our capability machine with call and return instructions and
well-bracketed control flow built-in to the operational semantics, and then
prove that compiling such code to our machine using our calling convention is
fully abstract~\citep{abadi_protection_1998}.

@@

Our logical relation is a unary, step-indexed Kripke logical relation
with recursive
worlds~\citep{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
Fundamental Theorem is similar to theirs and expresses
capability safety of the capability machine. Because we are not interested in
externally observable side-effects (like console output or memory access
traces), we do not require their notion of effect parametricity. Our logical
relation uses several ideas from previous work, like Kripke worlds with regions
containing state transition systems~\citep{Ahmed:popl09}, public/private future
worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

@@

\bibliographystyle{IEEEtranN}
% \bibliographystyle{plainnat}
%\bibliographystyle{plain}
\bibliography{IEEEabrv,references}
